### 2021-12-14 thesis/tex/ICtheory.tex
updated, and added worse s24 fig
update

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \input{commands}
% \begin{document}

@@

So how do we then obtain $\sigma_{x^r}(x^t)$ and $\mu(x^t)$ needed to construct the Gaussian? A Monte Carlo sample publicly released by the 
collaboration has all the ingredients that we need~\cite{IC2016}. In Table.~\ref{table:IC_MC} we show a selection of the data.
The "pdg" column refers to the Monte Carlo particle classification, where 13 is the tag for $\nm$, while -13 refers
to an $\anm$. Here we note a crucial property of the IceCube dataset that will impact our analysis: the MC released by the collaboration
only includes simulated muon events.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.4]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}.}\label{fig:IC_data}
\end{figure}

@@

We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}.
\begin{figure}[ht]
    \begin{centering}
    \includegraphics[scale=0.8]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as `Null' in red in the plots.}\label{fig:IC_MC_norm}
    \end{centering}
\end{figure} 
Thus, we are now able to sufficiently approximate the IceCube Monte Carlo, which makes us able to run simulations based on different physics scenarios.

>>>>

@@

So how do we then obtain $\sigma_{x^r}(x^t)$ and $\mu(x^t)$ needed to construct the Gaussian? A Monte Carlo sample publicly released by the 
collaboration has all the ingredients that we need~\cite{IC2016}. In Table.~\ref{table:IC_MC} we show a selection of the data.
The `pdg' column refers to the Monte Carlo particle classification, where 13 is the tag for $\nm$, while -13 refers
to an $\anm$. Here we note a crucial property of the IceCube dataset that will impact our analysis: the MC released by the collaboration
only includes simulated muon events.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.4]{figures/IC_data.pdf}
    \caption{IceCube track events from~\cite{IC2020}, displaying the $E^{-1.7}$ drop-off in event count.}\label{fig:IC_data}
\end{figure}

@@

We deemed these shapes to be satisfactory, thus allowing us to multiply Eq.~\ref{eq:ICevents} by the 
correction factors of Eq.~\ref{eq:MC_norm}.
\begin{figure}[ht]
    \begin{centering}
    \includegraphics[scale=0.7]{figures/IC_MC_norm.pdf}
    \caption{IceCube Monte Carlo, binned in $\Ereco$ and $\zreco$. We compare this with our simulations shown as `Null' in red in the plots.}\label{fig:IC_MC_norm}
    \end{centering}
\end{figure} 
Thus, we are now able to sufficiently approximate the IceCube Monte Carlo, which makes us able to run simulations based on different physics scenarios.

### 2021-12-14 thesis/tex/SM.tex
Lots of changes!!
update

% \input{commands}\begin{document}
\begin{figure}[!h]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/SM_figure.pdf}
\end{figure}
\section{The Standard Model}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight independent generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

@@

\subsection{Beyond the Standard Model}
Together, these three interactions make up the Standard Model gauge group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This determines the form of the three coupling constants, which numerical values must be experimentally measured. 
Since the vector bosons are represented by the generators, they are uniquely determined by the symmetry group. However, the scalar boson(s) and fermions are free as long as they
belong to representations of the symmetry group. By this construction, modifications to fermions rather than bosons are generally easier to make, allowing us to propose amendments to the model. Even the
number of fermions must be experimentally verified and can be altered from a phenomenological standpoint. In this work,
we will use this leniency to introduce a new particle and examine to which extent 
these modifications might be supported by experimental evidence. Moreover, we will examine the possibility of adding a new completely new force, which manifests itself as modifications to the neutrino-matter interactions.

@@

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa part of the interaction Lagrangian 
\begin{align} 
    \mathcal{L}_{int} \subseteq -\left( \frac{v + H}{\sqrt{2}} \right) \left[ \bar{\ell}^{\prime}_{\alpha L} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \bar{\nu}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,,
\end{align}
where $v$ is the Higgs vacuum expectation value, $H$ the Higgs field, $\ell_{\alpha L}^\prime (\ell_{\beta R}^\prime)$ the left (right) handed charged lepton field, and summation over the lepton indices $\alpha,\beta$ is implied. $Y_{\alpha \beta}^{\prime \ell}$ are the charged lepton Yukawa couplings, and $\nu^\prime_{\alpha L}(\nu^\prime_{\beta R})$ are the left (right) handed neutrino fields .
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, as emphasized with their prime. 
This can be remedied by diagonalizing $Y^{\prime \ell}$  with a biunitary transformation
\begin{align}
    V_{ L}^{\ell \dagger} Y^{\prime \ell} V_{ R}^{\ell}=Y^{\ell} \,.
\end{align}
where $V^\ell_L$ and $V^\ell_R$ are two unitary matrices.
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. We note that this will not yet generate neutrino masses, 
since no flavor rotation can make massless fields massive.
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.95\textwidth]{figures/SM_figure.pdf}
\end{figure}
\section{The Standard Model}\label{ch:SM}
In order to describe the three quantizable forces of nature, we gather the mediators of each force -- the vector bosons -- into local (gauge) symmetry groups. 
Each vector boson has one corresponding generator, the set of which constitutes the group.
The strong charge is mediated by eight massless gluons, which correspond to the eight independent generators of $\text{SU}(3)_C$. 
The weak charge is mediated by the three massive gauge bosons $W^\pm$ and $Z$, and the massless photon $\gamma$, 
which constitute the generators of $\text{SU}(2)_L$ and $\text{U}(1)_Y$.

@@

\subsection{Beyond the Standard Model}
Together, these three interactions make up the Standard Model gauge group $\mathrm{SU}(3)_{\mathrm{C}} \times \mathrm{SU}(2)_{L} \times \mathrm{U}(1)_{Y}$. 
This determines the form of the three coupling constants, which numerical values must be experimentally measured. 
Since the vector bosons are represented by the generators, they are uniquely determined by the symmetry group. However, the scalar boson(s) and fermions are free as long as they
belong to representations of the symmetry group. By this construction, modifications to fermions rather than bosons are generally easier to make, allowing us to propose amendments to the model. Even the
number of fermions must be experimentally verified and can be altered from a phenomenological standpoint. In this work,
we will use this leniency to introduce a new particle and examine to which extent 
these modifications might be supported by experimental evidence. Moreover, we will examine the possibility of adding a new completely new interaction, which manifests itself as modifications to the neutrino-matter interactions.

@@

We extend the Standard Model by adding a right-handed component of this field with neutrino Yukawa couplings $Y_{\alpha \beta}^{\prime \nu}$ to the Higgs-lepton Yukawa part of the interaction Lagrangian 
\begin{align} 
    \mathcal{L}_{int} \subseteq -\left( \frac{v + H}{\sqrt{2}} \right) \left[ \bar{\ell}^{\prime}_{\alpha L} Y_{\alpha \beta}^{\prime \ell} \ell_{\beta R}^{\prime} + \bar{\nu}_{\alpha L}^{\prime} Y_{\alpha \beta}^{\prime \nu} \nu_{\beta R}^{\prime}\right]\,,
\end{align}
where $v$ is the Higgs vacuum expectation value, $H$ the Higgs field, $\ell_{\alpha L}^\prime (\ell_{\beta R}^\prime)$ the left (right) handed charged lepton field, and summation over the lepton indices $\alpha,\beta$ is implied. $Y_{\alpha \beta}^{\prime \ell}$ are the charged lepton Yukawa couplings, and $\nu^\prime_{\alpha L}(\nu^\prime_{\beta R})$ are the left (right) handed neutrino fields .
Now, the Yukawa coupling matrices $Y^{\prime \ell}$ and $Y^{\prime \nu}$ are non-diagonal, as emphasized with their prime. 
This can be remedied by diagonalizing $Y^{\prime \ell}$  with a biunitary transformation
\begin{align}
    V_{ L}^{\ell \dagger} Y^{\prime \ell} V_{ R}^{\ell}=Y^{\ell} \,.
\end{align}
where $V^\ell_L$ and $V^\ell_R$ are two unitary matrices.
Similarly, we diagonalize and $Y^{\prime \nu}$ with a unitary matrix $V^\nu$. 
Now we state a crucial difference between the properties of the charged lepton and the neutrino fields.
While the charged lepton flavor eigenstate was uniquely determined by its mass eigenstate, the neutrino flavor is a superposition of mass eigenstates. 
This is because neutrinos are indirectly detected via the observation of its associated charged lepton, so there is no requirement for neutrino flavor eigenstates to have a definite mass. 
The flavor of a neutrino is then, by definition, the flavor of the associated charged lepton. 
This fact is denoted as giving the mass eigenstates Latin numerals and letters, while the flavor eigenstates stay as Greek letters.

### 2021-12-14 thesis/tex/introduction.tex
update

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we introduce a new particle: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we completely set the sterile neutrino aside and instead turn to the interactions between neutrinos and the Earth. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

Finally, Chapter~\ref{ch:results} contains the result from our two separate extensions of the Standard Model.
In Section~\ref{sec:sterileResults} we present our findings for the sterile neutrino and compare those with literature. 
In Section~\ref{sec:constraining}, we present new constrained bounds on the NSI parameters by combining our results for IceCube, DeepCore, and PINGU.

>>>>

Chapter~\ref{ch:theory} consists of two parts. In Section~\ref{sec:anomalies}, we consider a particle beyond the Standard Model: the sterile neutrino. 
We present how neutrino oscillations are modified by this hypothesized fourth neutrino and how these modifications would appear in IceCube if the particle is present in Nature. 
In Section~\ref{sec:nsiTheory}, we completely set the sterile neutrino aside and instead turn to the interactions between neutrinos and matter. 
We again amend the Standard Model by considering exotic interactions from a higher energy theory, manifesting themselves as sub-leading modifications to the matter potential. 
Since these new interactions are not present in the Standard Model, we call them non-standard interactions (NSI).

Finally, Chapter~\ref{ch:results} contains the result from our two separate extensions of the Standard Model.
In Section~\ref{sec:sterileResults} we present our findings for the sterile neutrino. 
In Section~\ref{sec:constraining}, we present new constrained bounds on the NSI parameters compare those with literature.

### 2021-12-14 thesis/tex/oscillations.tex
update

We construct the PMNS matrix by parametrizing it as as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12}P_{Maj} \nonumber \\
     &= \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
         U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
         U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
      \end{pmatrix} \nonumber \\
      & = \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix}
\begin{pmatrix}1 & 0 & 0 \\ 0 & e^{i \lambda_{2}} & 0 \\ 0 & 0 & e^{i \lambda_{3}}\end{pmatrix}
\nonumber \\
 &=\begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{12} c_{23}-s_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}}  & -c_{12} s_{23}-s_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{23} c_{13}
    \end{pmatrix}
    \begin{pmatrix}1 & 0 & 0 \\ 0 & e^{i \lambda_{2}} & 0 \\ 0 & 0 & e^{i \lambda_{3}}\end{pmatrix}\,,
\end{align}
where $s_{ij} = \sin(\theta_{ij})$, $c_{ij} = \cos(\theta_{ij})$, $0 \leq \theta_{ij} \leq \pi/2$ and $ 0 \leq \lambda_{i},\delta_\mathrm{CP} \leq 2\pi$.

By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure: one term in the form of a Dirac CP-violating phase $\delta_\mathrm{CP}$.
However, if the neutrinos are Majorana fermions, the rephasing of the left-handed fields would cause the neutrino masses to become complex.
Thus, we include two Majorana phases $\lambda_{2},\lambda_3$ in the PMNS matrix. These two phases are detectable in processes that violate lepton number, such as neutrinoless double $\beta$ decay.
Majorana phases are not detectable in neutrino oscillation experiments because oscillations are not violating lepton number, and a Majorana phase would
violate lepton number by two units.

@@

Experimental measurements show that $\theta_{13}$ is the smallest of the mixing angles~\cite{nufit}, causing 
$\nm$ to primarily mix into $\nt$. Using values from~\cite{nufit}, but taking $\delta_\mathrm{CP} = 0$ for simplicity,
the mixing matrix from~\ref{PMNS_def} takes the values
\begin{align}\label{eq:Uvalues}
   U = \begin{pmatrix}
       U_{e 1} & U_{e2} & U_{e3} \\
       U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
       U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
   \end{pmatrix} 
   = \begin{pmatrix}
       0.825 & 0.545 & 0.149 \\
       -0.455 & 0.485 & 0.746 \\
       0.334 & -0.684 & 0.649
   \end{pmatrix} \,,
\end{align}
which use to we plot the $P_{\mu \beta}$ probabilities in Fig.~\ref{fig:vac_osc}. Here, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12000}{\km}\footnote{This baseline
is approximately the Earth diameter, and will be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a suppressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
The smallness of $U_{e3}$ reduces the effect of $\dm[21]$ mass-squared difference,
allowing $\dm[31]$ to drive the oscillations. Thus, 
The most important mixing matrix elements for $P_{\a\b}$ are the ones appearing in the term with $\dm[31]$:
namely the product $U_{\a3}^*U_{\b3} U_{\b1}^*U_{\a1}$.

\begin{figure}[!h]
    \begin{center}
    \includegraphics[width=1\textwidth]{figures/vac_osc.pdf}
    \caption{Probabilities of $\nm$ oscillations after travelling 12 000 km in vacuum.}\label{fig:vac_osc}
    \end{center}
\end{figure}

@@

In Fig.~\ref{fig:vac_vs_matter}, we have plotted the oscillation probabilities for an incoming $\nm$ with \si{\GeV} energy,
traveling 12000 km. The red line is the vacuum oscillation, while the black line is the matter oscillation, traversing through the 
matter density profile of the Earth. We see that the matter effects are more for some energies, and less for others. At energies 
above $\SI{10}{\GeV}$, the matter effects are small for $\Pmm$ and $\Pmt$, while they heavily deviate for single-digit \si{\GeV} energies.
$\Pme$ behaves in a different way, where the matter effects affect the oscillation probability regardless of energy. Here, we start to
see the convergence as late as $\SI{100}{\GeV}$. Thus, matter effects are most important for energies below $\SI{100}{\GeV}$, for all channels.
Above these energies, the density profile of the Earth is less relevant, and the Earth is more or less transparent for $\nm$.

@@

% Let us return to the Hamiltonian with our two-neutrino mixing matrix. We have 
% \begin{align}
%     H_F
%         &= \frac{1}{2E}\left[R_{12} \begin{pmatrix}
%             0 & 0 & 0 \\
%             0 & \dm_{21} & 0 \\
%             0 & 0 & 0
%         \end{pmatrix} R_{12}^\dagger\right] + \sqrt{2}G_F N_e \begin{pmatrix}
%             1 & 0 & 0 \\
%             0 & 0 & 0 \\
%             0 & 0 & 0
%         \end{pmatrix} \nonumber \\
%         &= \frac{1}{2E}\left[\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1 \end{pmatrix}  \begin{pmatrix}
%             0 & 0 & 0 \\
%             0 & \dm_{21} & 0 \\
%             0 & 0 & 0
%         \end{pmatrix} \begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1 \end{pmatrix}^\dagger \right] + \sqrt{2}G_F N_e \begin{pmatrix}
%             1 & 0 & 0 \\
%             0 & 0 & 0 \\
%             0 & 0 & 0
%         \end{pmatrix}\,. 
% \end{align}
% TODO: segue into 2gen
We can reduce this down to two dimensions by extracting the upper-left block of all matrices. We now have the Hamiltonian in a two-dimensional form, and start simplifying:

@@

In Fig.~\ref{fig:2_vs_3}, we have plotted the oscillation probabilities for an incoming $\nm$ with \si{\GeV} energy,
traveling 12000 km. The black lines are the probabilities of $\nm$ oscillation, after it has traversing through the diameter of the Earth.
The red lines are the probabilities of $\nm$ oscillation, but in the two-neutrino picture using Eq.~\ref{eq:2gen}. So in the left-most panel,
we are only considering the mixing between $\ne$ and $\nm$, completely neglecting $\theta_{13}$ and $\theta_{23}$.

>>>>

We construct the PMNS matrix by parametrizing it with the rotation matrices $R_{ij}$ and a matrix $P_{Maj}$ as
\begin{align}\label{PMNS_def}
    U &= R_{23}R_{13,\delta_{\text{CP}}}R_{12}P_{Maj} \nonumber \\
     &= \begin{pmatrix}
        U_{e 1} & U_{e2} & U_{e3} \\
         U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
         U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
      \end{pmatrix} \nonumber \\
      & = \begin{pmatrix}1 & 0 & 0 \\ 0 & c_{23} & s_{23} \\ 0 & -s_{23} & c_{23}\end{pmatrix}
\begin{pmatrix}c_{13} & 0 & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 0 & 1 & 0 \\ -s_{13} e^{i \delta_{\mathrm{CP}}} & 0 & c_{13}\end{pmatrix}
\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1\end{pmatrix}
\begin{pmatrix}1 & 0 & 0 \\ 0 & e^{i \lambda_{2}} & 0 \\ 0 & 0 & e^{i \lambda_{3}}\end{pmatrix}
\nonumber \\
 &=\begin{pmatrix}c_{12} c_{13} & s_{12} c_{13} & s_{13} e^{-i \delta_{\mathrm{CP}}} \\ 
        -s_{12} c_{23}-c_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{12} c_{23}-s_{12} s_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & s_{23} c_{13} \\ 
        s_{12} s_{23}-c_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}}  & -c_{12} s_{23}-s_{12} c_{23} s_{13}e^{i \delta_{\mathrm{CP}}} & c_{23} c_{13}
    \end{pmatrix}
    \begin{pmatrix}1 & 0 & 0 \\ 0 & e^{i \lambda_{2}} & 0 \\ 0 & 0 & e^{i \lambda_{3}}\end{pmatrix}\,,
\end{align}
where $s_{ij} = \sin(\theta_{ij})$, $c_{ij} = \cos(\theta_{ij})$, $0 \leq \theta_{ij} \leq \pi/2$ and $ 0 \leq \lambda_{i},\delta_\mathrm{CP} \leq 2\pi$.

By construction, it provides the unitary transformation between the flavor and mass bases.
Any unitary $3\times3$ matrix can be parametrized using three angles and six phases. However, not all phases affect the 
charged and weak currents and are thus not observable by us. Moreover, both the Lagrangian and the currents are invariant under global $U(1)$ transformations,
leaving only one physical phase for us to measure: one term in the form of a Dirac CP-violating phase $\delta_\mathrm{CP}$.
However, if the neutrinos are Majorana fermions, the rephasing of the left-handed fields would cause the neutrino masses to become complex.
Thus, we include two Majorana phases $\lambda_{2},\lambda_3$ in the PMNS matrix through $P_{Maj}$. These two phases are detectable in processes that violate lepton number, such as neutrinoless double $\beta$ decay.
Majorana phases are not detectable in neutrino oscillation experiments because oscillations are not violating lepton number, and a Majorana phase would
violate lepton number by two units.

@@

Experimental measurements show that $\theta_{13}$ is the smallest of the mixing angles~\cite{nufit}, causing 
$\nm$ to primarily mix into $\nt$. Using values from~\cite{nufit}, but taking $\delta_\mathrm{CP} = 0$ for simplicity,
the mixing matrix from~\ref{PMNS_def} takes the values
\begin{align}\label{eq:Uvalues}
   U = \begin{pmatrix}
       U_{e 1} & U_{e2} & U_{e3} \\
       U_{\mu 1} & U_{\mu 2} & U_{\mu 3} \\
       U_{\tau 1} & U_{\tau 2} & U_{\tau 3}
   \end{pmatrix} 
   = \begin{pmatrix}
       0.825 & 0.545 & 0.149 \\
       -0.455 & 0.485 & 0.746 \\
       0.334 & -0.684 & 0.649
   \end{pmatrix} \,,
\end{align}
which use to we plot the $P_{\mu \beta}$ probabilities in Fig.~\ref{fig:vac_osc}. Here, we show the energy spectra of these oscillations in the low \si{\GeV} range for neutrinos that travel \SI{12742}{\km}\footnote{This baseline
is the Earth diameter, and will, apart from the atmosphere, be the maximum travel distance which we will use later on.}.
In the left-most panel, we see a suppressed $\nm \to \ne$ transition. We can understand this behavior by studying Eq.~\ref{eq:Pab} and the 
numerical values of the mixing matrix in Eq.~\ref{eq:Uvalues}.
The smallness of $U_{e3}$ reduces the effect of $\dm[21]$ mass-squared difference,
allowing $\dm[31]$ to drive the oscillations. Thus, 
The most important mixing matrix elements for $P_{\a\b}$ are the ones appearing in the term with $\dm[31]$:
namely the product $U_{\a3}^*U_{\b3} U_{\b1}^*U_{\a1}$.

\begin{figure}[!h]
    \begin{center}
    \includegraphics[width=1\textwidth]{figures/vac_osc.pdf}
    \caption{Probabilities of $\nm$ oscillations after travelling 12 742 km in vacuum, using parameters from~\cite{nufit}.}\label{fig:vac_osc}
    \end{center}
\end{figure}

@@

In Fig.~\ref{fig:vac_vs_matter}, we have plotted the oscillation probabilities for an incoming $\nm$ with \si{\GeV} energy,
traveling 12742 km. The red line is the vacuum oscillation, while the black line is the matter oscillation, traversing through the 
matter density profile of the Earth. We see that the matter effects are more for some energies, and less for others. At energies 
above $\SI{10}{\GeV}$, the matter effects are small for $\Pmm$ and $\Pmt$, while they heavily deviate for single-digit \si{\GeV} energies.
$\Pme$ behaves in a different way, where the matter effects affect the oscillation probability regardless of energy. Here, we start to
see the convergence as late as $\SI{100}{\GeV}$. Thus, matter effects are most important for energies below $\SI{100}{\GeV}$, for all channels.
Above these energies, the density profile of the Earth is less relevant, and the Earth is more or less transparent for $\nm$.

@@

% Let us return to the Hamiltonian with our two-neutrino mixing matrix. We have 
% \begin{align}
%     H_F
%         &= \frac{1}{2E}\left[R_{12} \begin{pmatrix}
%             0 & 0 & 0 \\
%             0 & \dm_{21} & 0 \\
%             0 & 0 & 0
%         \end{pmatrix} R_{12}^\dagger\right] + \sqrt{2}G_F N_e \begin{pmatrix}
%             1 & 0 & 0 \\
%             0 & 0 & 0 \\
%             0 & 0 & 0
%         \end{pmatrix} \nonumber \\
%         &= \frac{1}{2E}\left[\begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1 \end{pmatrix}  \begin{pmatrix}
%             0 & 0 & 0 \\
%             0 & \dm_{21} & 0 \\
%             0 & 0 & 0
%         \end{pmatrix} \begin{pmatrix}c_{12} & s_{12} & 0 \\ -s_{12} & c_{12} & 0 \\ 0 & 0 & 1 \end{pmatrix}^\dagger \right] + \sqrt{2}G_F N_e \begin{pmatrix}
%             1 & 0 & 0 \\
%             0 & 0 & 0 \\
%             0 & 0 & 0
%         \end{pmatrix}\,. 
% \end{align}
% TODO: segue into 2gen
We can reduce this down to two dimensions by extracting the upper-left block of all matrices\footnote{Or, following the method of the derivation of the Hamiltonianin the three-
neutrino picture, take U as one two-dimensional rotation matrix.}. We now have the Hamiltonian in a two-dimensional form, and start simplifying:

@@

In Fig.~\ref{fig:2_vs_3}, we have plotted the oscillation probabilities for an incoming $\nm$ with \si{\GeV} energy,
traveling through the entire Earth diameter. The black lines are the probabilities of $\nm$ oscillation, after it has traversing through the diameter of the Earth.
The red lines are the probabilities of $\nm$ oscillation, but in the two-neutrino picture using Eq.~\ref{eq:2gen}. So in the left-most panel,
we are only considering the mixing between $\ne$ and $\nm$, completely neglecting $\theta_{13}$ and $\theta_{23}$.
